---
title: "Mango"
author: "Erika Luna"
date: "21/10/2020"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width='950px', dpi=300)
```

This document explores the mango data from SIAP - Mexico.
- Data has been summarized at the state level.
- 26 states report mango production during the time period of 1980 - 2016.
- There are 962 observations total (26 states x 37 years), one observation/year for each state. 

```{r packages, include=FALSE}
#Load packages
library(tidyverse)
library(tidyr)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(DT)
library(forecast)
```

```{r load data}
SIAP <- read.csv("/Users/erikaluna/R\ Studio/msc_thesis/SIAP.csv") 
```

## Production cycle
```{r}
mango_agyear <- read.csv("mango_agyear.csv")
mango_agyear %>% 
  ggplot(aes(x=factor(Month,levels=month.name), Production)) +
  geom_bar(stat = "identity") +
  ylab("Production (%)") +
  xlab("Month") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust= 0.5))
```
## Water Supply
# Irrigated
```{r irrigated}
mango_irri_state <- SIAP %>% 
  filter(#state %in% c("baja california","baja california sur", "chihuahua", "durango", "nayarit", "sinaloa", "sonora"), 
    year < 1999, 
    crop == "mango", 
    water == "irrigated", 
    production > 0) %>% 
  select(year, state, water, yield, production, harvested)
#mango_irri_state %>% 
  #DT::datatable()

mango_irri_mun <- SIAP %>% 
  filter(#state %in% c("baja california","baja california sur", "chihuahua", "durango", "nayarit", "sinaloa", "sonora"), 
    year > 1998, 
    crop == "mango", 
    water == "irrigated", 
    production > 0) %>% 
  select(year, state, water, yield, production, harvested)
#mango_irri_state %>% 
  #DT::datatable()

### Exploring the irrigated and rainfed data: 

ggplot(mango_irri_state, aes(fill=water, x=year)) + 
    geom_bar(stat = "count") +
  facet_wrap(~state)


SIAP %>%  #in theory, data is at the municipal level starting on 2003. But more than two observations are present starting on 1999. Municipal id start on 2003. 
  filter(state == "puebla") %>% 
  filter(crop == "mango") %>% 
  filter(year == 2004) %>% 
  DT::datatable()
```
# Rainfed
```{r rainfed}
mango_rain_state <- SIAP %>% 
  filter(#state %in% c("baja california","baja california sur", "chihuahua", "durango", "nayarit", "sinaloa", "sonora"), 
         year < 1999, 
         crop == "mango", 
         water == "rainfed", 
         production > 0) %>% 
  select(year, state, water, yield, production, harvested) 
#mango_rain_state %>% 
  #DT::datatable()
mango_rain_state %>% 
  ggplot(aes(water, yield)) +
  geom_boxplot() +
  scale_y_continuous("Yield (ha)", trans = "log10") +
  ylab("Yield (ha)") +
  xlab("Water supply")

mango_rain_mun <- SIAP %>% 
  filter(#state %in% c("baja california","baja california sur", "chihuahua", "durango", "nayarit", "sinaloa", "sonora"), 
         year > 1998, 
         crop == "mango", 
         water == "rainfed",  
         production > 0) %>% 
  select(year, state, water, yield, production, harvested)
#mango_rain_mun %>% 
  #DT::datatable()
mango_rain_mun %>% 
  ggplot(aes(water, yield)) +
  geom_boxplot() +
  scale_y_continuous("Yield (ha)", trans = "log10") +
  ylab("Yield (ha)") +
  xlab("Water supply")
```

# Irrigated and Rainfed
```{r}
mango_water_state <- bind_rows(mango_irri_state, mango_rain_state)
mango_water_mun <- bind_rows(mango_irri_mun, mango_rain_mun)
```

```{r}
mango_water_state %>% 
  #DT::datatable()
  
  
  ggplot(aes(water, yield)) +
  geom_boxplot() +
  scale_y_continuous("Yield (ha)", trans = "log10") +
  ylab("Yield (ha)") +
  xlab("Water mode") +
  ggtitle("Mango State level")
  #scale_y_continuous(labels = comma) +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust= 0.5))

mango_all <- mango_water_mun%>% 
  ggplot(aes(year, production, color = water)) + 
  geom_point() +
  #geom_smooth(method = "lm", se = T) +
  ylab("Yields (t/ha)") +
  xlab("Years") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
  #geom_rect(data = subset(mango, state %in% c(mango_complete$state)), 
            #fill = NA, colour = "red", xmin = -Inf,xmax = Inf,
            #ymin = -Inf,ymax = Inf) +     
  facet_wrap(~state, scales="free_y", ncol=5)
  #facet_wrap(~state, ncol=5)
mango_all

tab <- table(mango_water_state$water, mango_water_state$state)

prop.table(tab)

tab <- with(mango_water_state, table(water, state))
prop.table(tab, margin = 1)
prop.table(tab, 1)*100
tab <- as.data.frame.matrix(tab)
t <- data.frame(rbind(tab))
t<- t %>% 
  gather(state)
irr_rain <- rep(c("irrigated", "rainfed"), 32) 
t <- cbind(t, irr_rain)
t <- t %>% 
  group_by(state) %>% 
mutate(percent = 100*value/sum(value)) %>% 
  drop_na()

irr_states <- t %>% 
  filter(irr_rain == "irrigated")

write.csv(irr_states, "irr_states.csv")

rain_states <- t %>% 
  filter(irr_rain == "rainfed")

write.csv(rain_states, "rain_states.csv")

write.csv(t, "water_freq.csv")

```

```{r}
mango_water_mun %>% 
  ggplot(aes(water, yield)) +
  geom_boxplot() +
  scale_y_continuous("Yield (ha)", trans = "log10") +
  ylab("Yield (ha)") +
  xlab("Water mode") +
  ggtitle("Mango Municipal level")
  #scale_y_continuous(labels = comma) +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust= 0.5))
```

## Mango
```{r crop}
one_crop <- SIAP %>% 
  filter(crop == "mango") %>% 
  group_by(year, state) %>% 
  summarise(ag_yield = round(sum(production)/sum(harvested), digits = 2),
            ag_prod = sum(production),
            ag_planted = sum(planted),
            ag_harv = sum(harvested), 
            ag_losses = sum(losses))
```
## Mango national
```{r}
mango_nat <- SIAP %>% 
  filter(crop == "mango") %>% 
  #filter(state %in% c("baja california sur", "durango", "nayarit", "sinaloa", "sonora"), crop == "mango") %>% 
  group_by(year) %>% 
  summarise(ag_yield = round(sum(production)/sum(harvested), digits = 2),
            ag_prod = sum(production),
            ag_planted = sum(planted),
            ag_harv = sum(harvested), 
            ag_losses = sum(losses))

mango_nat %>% 
  DT::datatable()
```

## Number of observations
```{r NAs, echo=FALSE}
crop_full <- one_crop %>% 
  group_by(state) %>% 
  summarise(obs = sum(!is.na(ag_prod))) 

plot_obs <- crop_full %>% 
  ggplot(aes(state, obs)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_y_continuous(name = "number of observations") #+
  #ggtitle("NAs for production")
plot_obs
```





## Panel data
A data frama with all years and all states that grow mango.
```{r Panel}
period <- tibble(rep(c(1980:2016), times = 26)) #26 states report mango production
colnames(period) <- c("year") 
states <- tibble(rep(c("baja california", "baja california sur", "campeche", 
                       "chiapas", "colima", "durango",
                        "guanajuato", "guerrero", "hidalgo", "jalisco", 
                       "mexico", "michoacan", "morelos", "nayarit", "oaxaca", 
                       "puebla", "queretaro", "quintana roo",
                        "san luis potosi", "sinaloa", "sonora", "tabasco", 
                       "tamaulipas", "veracruz", "yucatan", "zacatecas"), times = 37))
colnames(states) <- c("state") 
states <- states %>% 
  arrange(state)
states_period <- cbind(states, period)
```


## Data frame for Mango
```{r Mango}
mango <- left_join(states_period, one_crop, by=c("state", "year"))
mango <- mango %>%  
  transform(i=as.numeric(factor(state))) %>% 
  transform(t=as.numeric(factor(year))) %>% 
  group_by(year) %>% 
  arrange(state) 

mango %>% 
  DT::datatable()
```


## Plots
### Production

```{r}
mango %>% 
  group_by(state) %>% 
  summarise(max_prod = max(ag_prod, na.rm=T),
            min_prod = min(ag_prod, na.rm=T),
            range_prod = max(ag_prod, na.rm=T) - min(ag_prod, na.rm=T),
            sd_prod = sd(ag_prod, na.rm=T),
            mean_prod = mean(ag_prod, na.rm=T),
            median_prod = median(ag_prod, na.rm=T)) %>% 
  knitr::kable()
```

```{r}
mango %>% 
  ggplot(aes(state, ag_prod)) +
  geom_boxplot() +
  ylab("Production (tonnes)") +
  xlab("State") +
  #scale_y_continuous(labels = comma) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust= 0.5))
```



```{r}
number_obs <- mango %>% 
  group_by(state) %>% 
  summarise(obs = sum(!is.na(ag_prod)))

mango_complete <- number_obs %>% 
  filter(obs > 34)
mango_complete

mango_ts <- mango %>% 
  ggplot(aes(year, ag_prod)) + 
  geom_line()+
  ylab("Production (tonnes)") +
  xlab("Years") +
  ggtitle("Mango Production 1980 - 2016") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
  geom_rect(data = subset(mango, state %in% c(mango_complete$state)), 
            fill = NA, colour = "red", xmin = -Inf,xmax = Inf,
            ymin = -Inf,ymax = Inf) +     
  facet_wrap(~state, scales="free_y", ncol=5) 
  #facet_wrap(~state, ncol=5)
mango_ts  
```

### Yield
```{r}
mango %>% 
  group_by(state) %>% 
  summarise(max_yield = max(ag_yield, na.rm=T),
            min_yield = min(ag_yield, na.rm=T),
            range_yield = max(ag_yield, na.rm=T) - min(ag_yield, na.rm=T),
            sd_yield = sd(ag_yield, na.rm=T),
            mean_yield = mean(ag_yield, na.rm=T),
            median_yield = median(ag_yield, na.rm=T)) %>% 
  knitr::kable()
```

```{r}
mango %>% 
  filter(state %in% c("baja california sur", "durango", "sinaloa", "sonora")) %>% 
  ggplot(aes(state, ag_yield)) +
  geom_boxplot() +
  ylab("Yield (tonnes/ha)") +
  xlab("State") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust= 0.5))
```



```{r}
number_obs <- mango %>% 
  group_by(state) %>% 
  summarise(obs = sum(!is.na(ag_yield)))

mango_complete <- number_obs %>% 
  filter(obs > 34)
mango_complete

mango_ts <- mango %>% 
  ggplot(aes(year, ag_yield)) + 
  geom_line()+
  ylab("Yield (tonnes/ha)") +
  xlab("Years") +
  ggtitle("Mango Yields 1980 - 2016") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
  geom_rect(data = subset(mango, state %in% c(mango_complete$state)), 
            fill = NA, colour = "red", xmin = -Inf,xmax = Inf,
            ymin = -Inf,ymax = Inf) +     
  facet_wrap(~state, scales="free_y", ncol=5) 
  #facet_wrap(~state, ncol=5)
mango_ts  
```

### Area
```{r}
mango %>% 
  group_by(state) %>% 
  summarise(max_area = max(ag_harv, na.rm=T),
            min_area = min(ag_harv, na.rm=T),
            range_area = max(ag_harv, na.rm=T) - min(ag_harv, na.rm=T),
            sd_area = sd(ag_harv, na.rm=T),
            mean_area = mean(ag_harv, na.rm=T),
            median_area = median(ag_harv, na.rm=T)) %>% 
  knitr::kable()
```

```{r}
mango %>% 
  ggplot(aes(state, ag_harv)) +
  geom_boxplot() +
  ylab("Area (tonnes)") +
  xlab("State") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust= 0.5))
```



```{r}
number_obs <- mango %>% 
  group_by(state) %>% 
  summarise(obs = sum(!is.na(ag_harv)))

mango_complete <- number_obs %>% 
  filter(obs > 34)
mango_complete

mango_ts <- mango %>% 
  ggplot(aes(year, ag_harv)) + 
  geom_line()+
  ylab("Area harvested (ha)") +
  xlab("Years") +
  ggtitle("Mango - Area Harvested (ha) 1980 - 2016") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
  geom_rect(data = subset(mango, state %in% c(mango_complete$state)), 
            fill = NA, colour = "red", xmin = -Inf,xmax = Inf,
            ymin = -Inf,ymax = Inf) +     
  facet_wrap(~state, scales="free_y", ncol=5) 
  #facet_wrap(~state, ncol=5)
mango_ts  
```




## Yields ~ Time
# Mango trends at the national level
```{r country - yields - years}
m <- mango_nat %>% 
  ggplot(aes(year, ag_yield)) + 
  geom_point() +
  geom_smooth(method = "lm", se = T) +
  stat_cor()+
  #scale_y_continuous(trans = 'log10')+
  #annotation_logticks(sides="lb")+
  ylab("Yield (tonnes/ha)") +
  xlab("Years") +
  ggtitle("MANGO Yields from 1980 to 2016")# +
  #theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))
m
```
# Mango trends at the state level
```{r state - yields - years}
lt <- mango %>% 
  filter(state %in% c("baja california sur", "durango", "sinaloa", "sonora")) %>% 
  ggplot(aes(year, ag_harv, color = state)) + 
  geom_line() +
  geom_smooth(method = "lm", se = T) +
  ylab("Yields (tonnes/ha)") +
  xlab("Years") +
  ggtitle("Mango Yields") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) #+
  #geom_rect(data = subset(mango, state %in% c(mango_complete$state)), 
            #fill = NA, colour = "red", xmin = -Inf,xmax = Inf,
            #ymin = -Inf,ymax = Inf) +     
  #facet_wrap(~state, scales="free_y", ncol=5) 
lt
#ggsave()
```



## Yields ~ Temperature
```{r load temp}
tmax_summary <- read.csv("/Users/erikaluna/R\ Studio/climate_data/tmax_summary.csv")
tmin_summary <- read.csv("/Users/erikaluna/R\ Studio/climate_data/tmin_summary.csv")
```

```{r mango - temp}
mango_tmax <- left_join(mango, tmax_summary, by = c("state" = "state", "year" = "year"))
mango_tmin <- left_join(mango, tmin_summary, by = c("state" = "state", "year" = "year"))
```

# Mango yields ~ temperature at the national level

```{r country - yields - tmax}
yt <- mango_tmax %>% 
  ggplot(aes(mean_tmax, ag_yield)) + 
  geom_point() +
  geom_smooth(method = "lm", se = T) +
  ylab("Yield (tonnes/ha)") +
  xlab("Temperature (ºC) ") +
  ggtitle("Mango Yields ~ Max Temperature")
yt
```

```{r country - yields - tmin}
yt <- mango_tmin %>% 
  ggplot(aes(mean_tmin, ag_yield)) + 
  geom_point() +
  geom_smooth(method = "lm", se = T) +
  ylab("Yield (tonnes/ha)") +
  xlab("Temperature (ºC) ") +
  ggtitle("Mango Yields ~ Min Temperature") #+
yt
```


```{r states - yields - tmax}
mango_lm <- mango_tmax %>% 
  filter(state %in% c("baja california sur", "durango", "sinaloa", "sonora")) %>% 
  #filter(state %in% c("baja california sur", "durango", "nayarit", "sinaloa", "sonora")) %>% 
  ggplot(aes(mean_tmax, ag_yield)) + 
  geom_point() +
  geom_smooth(method = "lm", se = T) +
  #scale_y_continuous(trans = 'log10')+
  ggtitle("mango Yields") +
   ylab("Yield (tonnes/ha)") +
  xlab("Temperature (ºC) ") +
  ggtitle("Mango Yields ~ Max Temperature") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
  facet_wrap(~state, scales="free_y") 
mango_lm
```
```{r states - yields - tmin}
mango_lm <- mango_tmin %>% 
  filter(state %in% c("baja california sur", "durango", "nayarit", "sinaloa", "sonora")) %>% 
  ggplot(aes(mean_tmin, ag_yield)) + 
  geom_point() +
  geom_smooth(method = "lm", se = T) +
  ylab("Yield (tonnes/ha)") +
  xlab("Temperature (ºC)") +
  #scale_y_continuous(trans = 'log10')+
  ggtitle("Mango Yields ~ Min Temperature") +
  facet_wrap(~state, scales="free_y") 
mango_lm
```

```{r}

```
## lm state level
```{r}
#### Packages ####
library(broom)

#### Prepare data ####
mango_states <- SIAP %>%
  filter(crop == "mango") %>% 
  drop_na() %>% 
  group_by(state_code, year) %>% #we group by state code so we can merge with the shp file 
  summarise(ag_yield = round(sum(production)/sum(harvested), digits = 2), 
            ag_prod = sum(production),
            ag_harv = sum(harvested))

mango_states %>% 
  DT::datatable()

#### Yields ####
#fitted_yields = mango_states %>% 
  #group_by(state_code) %>% 
  #do(model = rlm(ag_yield ~ year-1, data = ., na.action=na.exclude))

fitted_yields = mango_states %>%  
  group_by(state_code) %>% 
  #do(model = lm(scale(ag_yield) ~ scale(year), data = ., na.action=na.exclude))
  do(model = lm(ag_yield ~ year, data = ., na.action=na.exclude))
  
fitted_yields <- fitted_yields %>% tidy(model) %>% as_tibble()

mango_lm_yield <- fitted_yields %>% 
  #filter(term == "scale(year)") %>% 
  filter(term == "year") #%>% 
  #filter(p.value < 0.05) %>% 
  #mutate(two_se = 2*std.error) #%>% 
  #drop_na() %>% 
  #filter(abs(estimate) > two_se)

mango_lm_yield %>% 
  DT::datatable()

library("robustbase")
mango_states %>% 
  group_by(state_code) %>% 
  lmrob(ag_yield ~ year)


# lm(scale(your.y) ~ scale(your.x), data=your.Data)


#### Production ####
#fitted_prod = mango_states %>% 
 # group_by(state_code) %>% 
  #do(model = lm(ag_prod ~ year-1, data = ., na.action=na.exclude))
fitted_prod = mango_states %>%  
  group_by(state_code) %>% 
  #do(model = lm(scale(ag_prod) ~ scale(year), data = ., na.action=na.exclude))
  do(model = lm(ag_prod ~ year, data = ., na.action=na.exclude))

fitted_prod <- fitted_prod %>% tidy(model) %>% as_tibble()

mango_lm_prod <- fitted_prod %>% 
  #filter(term == "scale(year)") %>% 
  filter(term == "year")# %>% 
  #filter(p.value < 0.05) %>% 
  #mutate(two_se = 2*std.error) #%>% 
  #drop_na() %>% 
  #filter(abs(estimate) > two_se)




mango_lm_prod %>% 
  DT::datatable()

#### Harvested Area ####
#fitted_area = mango_states %>% 
  #group_by(state_code) %>% 
  #do(model = lm(ag_harv ~ year-1, data = ., na.action=na.exclude))

fitted_area = mango_states %>%  
  group_by(state_code) %>% 
  do(model = lm(ag_harv ~ year, data = ., na.action=na.exclude))
  
fitted_area <- fitted_area %>% tidy(model) %>% as_tibble()

mango_lm_area <- fitted_area %>% 
  filter(term == "year") #%>% 
  #filter(p.value < 0.05) %>% 
  #mutate(two_se = 2*std.error) #%>% 
  #drop_na() %>% 
  #filter(abs(estimate) > two_se)

mango_lm_area %>% 
  DT::datatable()


#### Write files ####
write.csv(mango_lm_yield, file = "mango_lm_yield.csv")
write.csv(mango_lm_prod, file = "mango_lm_prod.csv")
write.csv(mango_lm_area, file = "mango_lm_area.csv")
```

## lm mun level
```{r}

mango_mun <- SIAP %>% 
  filter(crop == "mango") %>% 
  filter(year > 2002) %>% 
  group_by(mun_code, year) %>% 
  summarise(ag_yield = round(sum(production)/sum(harvested), digits = 2),
            ag_prod = sum(production),
            ag_harv = sum(harvested))

tmp <- mango_mun %>% 
  drop_na() %>% 
  filter_all(all_vars(!is.infinite(.)))


fitted_models = tmp %>% 
  group_by(mun_code) %>% 
  do(model = lm(ag_yield ~ year, data = .))
fitted_models
library(broom)
v <- fitted_models %>% tidy(model) %>% as_tibble()
v %>% 
  DT::datatable()

mango_lm_mun <- v %>% 
  filter(term == "year") %>% 
  #filter(p.value < 0.05) %>% 
  mutate(two_se = 2*std.error) %>% 
  drop_na() %>% 
  filter(abs(estimate) > two_se)
  
  #summarise(zia = 2*std.error)

write.csv(mango_lm_mun, file = "mango_lm_mun.csv")
```

```{r}
mango_water <- SIAP %>% 
  filter(crop == "mango") %>% 
  drop_na() %>% 
  group_by(water, year) %>% 
  summarise(ag_yield = round(sum(production)/sum(harvested), digits = 2),
              ag_prod = sum(production),
              ag_harv = sum(harvested))

mango_water %>% 
    #ggplot(aes(x=year, y=ag_yield, color = water)) +
    #ggplot(aes(x=year, y=ag_prod, color = water)) +
    ggplot(aes(x=year, y=ag_harv, color = water)) +
    #scale_y_continuous("Yield (t/ha)") +
    #scale_y_continuous("Production (t)", labels = scales::comma) +
    scale_y_continuous("Harvested Area (ha)", labels = scales::comma) +
    geom_line() +
    geom_smooth(method = MASS::rlm, se = F, size = 0.5) +
    stat_cor() +
    labs(x="Year")

#ggsave("mango_water_yields.png")
#ggsave("mango_water_prod.png")
#ggsave("mango_water_harv.png")
    #facet_wrap(~state, scales="free_y")
```

```{r echo=TRUE}
SIAP %>% 
  filter(crop == "mango") %>% 
  drop_na() %>% 
  group_by(year) %>% 
  summarise(ag_yield = round(sum(production)/sum(harvested), digits = 2),
              ag_prod = sum(production),
              ag_harv = sum(harvested)) %>% 
    ggplot(aes(x=year, y=ag_yield)) +
    scale_y_continuous("Yield (t/ha)") +
    geom_line() +
    geom_smooth(method = MASS::rlm, se = F) +
    stat_cor()#+
    #facet_wrap(~state, scales="free_y")
```

```{r}
tmp %>% 
  ggplot(aes(year, ag_prod)) +
  geom_line() +
  geom_line(aes(y = ag_harv), color = "red") +
  scale_y_continuous("Production (tonnes)", trans = "log10", labels = scales::comma,
                     sec.axis = sec_axis(~., name = "Harvested Area (ha) ", labels = scales::comma)) +
  scale_x_continuous("Year") +
  ggtitle("Mango Production and Harvested Area") #+
  #facet_wrap(~state, scales="free_y") 

#ylim.prim <- c(0, 400000)   # in this example, precipitation
#ylim.sec <- c(0, 40000)    # in this example, temperature
ylim.prim <- c(0, 2000000)   # in this example, precipitation
ylim.sec <- c(0, 200000)

b <- diff(ylim.prim)/diff(ylim.sec)
a <- b*(ylim.prim[1] - ylim.sec[1])

SIAP %>% 
  filter(crop == "mango") %>% 
  drop_na() %>% 
  group_by(year) %>% 
  summarise(ag_yield = round(sum(production)/sum(harvested), digits = 2),
              ag_prod = sum(production),
              ag_harv = sum(harvested)) %>%
  #DT::datatable()
  ggplot(aes(year, ag_prod)) +
  geom_line() +
  geom_line(aes(y = ag_harv), color = "red") +
  scale_y_continuous("Production (tonnes)", trans = "log10", labels = scales::comma,
                     sec.axis = sec_axis(~ (. - a)/b, name = "Harvested Area (ha) ", labels = scales::comma)) +
  scale_x_continuous("Year") +
  ggtitle("Mango Production and Harvested Area") #+
  #facet_wrap(~state, scales="free_y") 
```

```{r}
mango_irrigated <- SIAP %>% 
  filter(crop == "mango") %>% 
  filter(water == "irrigated") %>% 
  filter(state %in% c("baja california sur", "durango", "sinaloa", "sonora")) %>% 
  #drop_na() %>% 
  group_by(state, year) %>% 
  summarise(ag_yield = round(sum(production)/sum(harvested), digits = 2),
              ag_prod = sum(production),
              ag_harv = sum(harvested))

mango_rainfed <- SIAP %>% 
  filter(crop == "mango") %>% 
  filter(water == "rainfed") %>% 
  filter(state %in% c("baja california sur", "durango", "sinaloa", "sonora")) %>% 
  #drop_na() %>% 
  group_by(state, year) %>% 
  summarise(ag_yield = round(sum(production)/sum(harvested), digits = 2),
              ag_prod = sum(production),
              ag_harv = sum(harvested))
```



## log(yield) - log(trend)
```{r}
#Extract the yields variable
mango_irrigated_yields <- mango_irrigated$ag_yield
mango_rainfed_yields <- mango_rainfed$ag_yield

#Convert to time series object
mango_irrigated_yields_ts <- ts(mango_irrigated_yields, start = 1980, frequency = 4)
mango_rainfed_yields_ts <- ts(mango_rainfed_yields, start = 1980, frequency = 4)

#NA
mango_irrigated_yields_ts_nona <- na.approx(mango_irrigated_yields_ts)
mango_rainfed_yields_ts_nona <- na.approx(mango_rainfed_yields_ts)

#Obtain trend
mango_irrigated_yields_trend <- ma(mango_irrigated_yields_ts_nona, order = 3, centre = T )
mango_rainfed_yields_trend <- ma(mango_rainfed_yields_ts_nona, order = 3, centre = T )

#Remove trend by applying log to ts and trend
mango_irrigated_detrend <- log(mango_irrigated_yields_ts_nona) - log(mango_irrigated_yields_trend)
mango_rainfed_detrend <- log(mango_rainfed_yields_ts_nona) - log(mango_rainfed_yields_trend)

#Plot
log_yield_log_trend <- cbind(mango_irrigated_detrend, mango_rainfed_detrend)
autoplot(log_yield_log_trend)
```
```{r}
library("TTR")
library("zoo")

tmp <- na.approx(mango_irrigated_yields_ts)
ma <- ma(tmp, order = 3, centre = F )
plot()
plot(ma, col = "red")

plot(wineind)
sm <- ma(wineind,order=12)
plot(sm,col="red")


```



```{r}
tmax_ts <- tmax_summary %>% 
  group_by(year) %>% 
  summarise(mean_ts = mean(mean_tmax, na.rm = T))

log_yield_log_trend_df <- as.tibble(log_yield_log_trend)

mango_log_tempmax <- cbind(log_yield_log_trend_df, tmax_ts)

mango_log_tempmax %>% 
  ggplot(aes(mean_ts, mango_irrigated_detrend)) + 
  geom_point() +
  #geom_smooth(method = "lm", se = T) +
  ylab("log (yield) - log (trend)") +
  xlab("Temperature (ºC) ") #+
  #ggtitle("Mango Yields ~ Max Temperature")
ggsave("mango_log_max.png")

max <- tmax_summary %>% 
  filter(state %in% c("baja california sur", "durango", "sinaloa", "sonora")) %>% 
  filter(year != 1985)

min <- tmin_summary %>% 
  filter(state %in% c("baja california sur", "durango", "sinaloa", "sonora")) %>% 
  select(mean_tmin)

temp_summary <- cbind(max, min)
temp_summary <- gather(temp_summary, "temp", "value", mean_tmax, mean_tmin)

temp_summary %>% 
  #group_by(state, year) %>% 
  ggplot(aes(year, value, color = temp)) + 
  geom_line() +
  #geom_smooth(method = "lm", se = T) +
  ylab("Temperature (ºC)") +
  xlab("Year") +
  theme(legend.title=element_blank()) +
  scale_color_manual(labels = c("Max", "Min"), values = c("red", "blue")) +
  #ggtitle("Temperature") +
  facet_wrap(~state)



ggsave("temperature.png")

```
```{r echo=FALSE}
mango_all <- SIAP %>%
#t<-    SIAP %>%
  filter(crop == "mango") %>% 
  #filter(state %in% c("baja california sur", "durango", "sinaloa", "sonora")) %>% 
  ggplot(aes(year, harvested, color = water)) + 
  geom_point() +
  #geom_smooth(method = "lm", se = T) +
  ylab("Yields (t/ha)") +
  xlab("Years") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
  #geom_rect(data = subset(mango, state %in% c(mango_complete$state)), 
            #fill = NA, colour = "red", xmin = -Inf,xmax = Inf,
            #ymin = -Inf,ymax = Inf) +     
  facet_wrap(~state, scales="free_y", ncol=5) 
mango_all



  mango_all <- SIAP %>%
  filter(crop == "mango")# %>% 
  filter(state %in% c("baja california sur", "durango", "sinaloa", "sonora")) %>% 
  ggplot(aes(year, harvested, color = state)) + 
  geom_point() +
  #geom_smooth(method = "lm", se = T) +
  ylab("Yields (t/ha)") +
  xlab("Years") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) #+
  #geom_rect(data = subset(mango, state %in% c(mango_complete$state)), 
            #fill = NA, colour = "red", xmin = -Inf,xmax = Inf,
            #ymin = -Inf,ymax = Inf) +     
  #facet_wrap(~state, scales="free_y", ncol=5) 
ggsave("state_yields_all.png")
ggsave("state_yields.png")
write.csv(mango_all, "mango.csv")
```

```{r}

water <- SIAP %>%
  filter(crop == "mango")
         #state == "durango") 

table(water$water)
table3 <- table(water$water, water$state)
prop.table(table3)

# re-order levels
reorder_size <- function(x) {
        factor(x, levels = names(sort(table(x), decreasing = TRUE)))
}

ggplot(mango_all, aes(x = reorder_size(state))) +
        geom_bar() +
        xlab("State") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(water, aes(x = reorder_size(state))) +
        geom_bar(aes(y = (..count..)/sum(..count..))) +
        xlab("State or Province") +
        scale_y_continuous(labels = scales::percent, name = "Proportion") +
        facet_grid(~ water) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(data = water, aes(x = state, y = harvested, fill = water)) + 
  geom_bar(stat = "identity", position = position_dodge()) + 
  xlab("Frequency of Father Visits") + ylab("Percent of Children")

d <- ggplot(water, aes(x= state,  group=water)) + 
    geom_bar(aes(y = ..prop.., stat="count")) +
    geom_text(aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -.5) +
    labs(y = "Percent") +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust= 0.5)) +
    facet_grid(~water)# +
    scale_y_continuous(labels = scales::percent)

# Stacked + percent
q <- mango_all %>% 
  ggplot(aes(fill=water, y=yield, x=reorder(state, yield))) + 
    geom_bar(position="fill", stat="identity") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust= 0.5))

mango_all %>% 
  group_by(state) %>% 
  summarise(mean_prod = mean(production)) %>% 
  ggplot(mango_all, aes(fill=water, y=mean_prod, x=state)) + 
    geom_bar(position="stack", stat="identity") +
   scale_y_continuous("Harvested Area (ha)", labels = scales::comma) +
    xlab("")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust= 0.5))
q

q <- ggplot(mango_all, aes(x=reorder(state, -production, mean), y=harvested, fill=water)) + 
   #geom_bar(position="stack", stat="identity") +
  geom_bar(stat="identity")+
   #scale_y_continuous("Production (t)", labels = scales::comma) +
  scale_y_continuous("Harvested area (ha)", labels = scales::comma) +
   # xlab("")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust= 0.5))


p=ggplot(data=mango_all)  
p+geom_bar(stat="identity")+  #
aes(x=reorder(state,-production,sum),y=production,label=production,fill=water)+
theme()
```




